---
title: "Ozarks FA Analysis"
output: html_document
date: "2022-10-28"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "./")
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
library(ggfortify)
#remotes::install_github("gavinsimpson/ggvegan")
#library(ggvegan)
library(here)
library(broom)
library(heplots)
library(broom.mixed)
library(car)
#library(leaps)

here::i_am("analysis/paper/Ozarks_FA_Analysis.Rmd")
```


# Setup
PUFAs: 
1) `18:2w6` = LIN 
2) `18:3w3` = ALA
3) `18:4w3` = SDA
4) `20:5w3` = EPA
5) `20:4w6` = ARA
6) `22:6w3` = DHA
For periphyton, we want to look at PUFA based on composition (ug PUFA/L) and surface area.
So need to create a compositional and surface area based periphyton matrix
For inverts, we selected taxa with FA measurements from at least three sites
```{r}
FA <- read_csv(here("analysis/data/derived_data/FA_all_tidied.csv"))|>
  select(where(~ any(. != 0)))

#invert units are ug FA/mg fresh weight
#periphyton units are ug FA/m2

FA2<-FA|>group_by(Site,EcoRegion)|>summarize(SRP_ug.L = mean(SRP_ug.L),NO3_ug.L = mean(NO3_ug.L), Canopy_Cover_per = mean(Canopy_Cover_per))
ggplot(data = FA2, aes(x = NO3_ug.L))+geom_histogram()
ggplot(data = FA2, aes(x = Canopy_Cover_per))+geom_density()


```



```{r prep FA data}
FA_groups_areal <- FA|>
  mutate(across(all_of(4:40), ~ifelse(Type == "Periphyton",
                                      ./1000, #periphyton units now mg FA/m2, leaves inverts alone
                                      .)))|> 
  rowwise()|>
  mutate(Tot_FA = sum(across(all_of(4:40))),
         MUFA = sum(c_across(contains(":1"))),
         MUFA_16 = sum(c_across(contains("16:1"))),
         MUFA_18 = sum(c_across(contains("18:1"))),
         SAFA = sum(c_across(c(matches(":0"),-matches("i-")))),
         LSAFA = sum(c_across(c(`22:0`, `24:0`,`23:0`))),
         brFA = sum(c_across(contains("i-"))),
         diatom_16PUFA = sum(c_across(c(`16:2w7`,`16:3w4`))),
         green_16PUFA = sum(c_across(c(`16:3w3`,`16:4w3`,`16:2w6`))),
         w3_sum = sum(c_across(contains("w3"))),
         w6_sum = sum(c_across(contains("w6"))),
         w3_w6 = w3_sum/w6_sum,
         EPA = `20:5w3`,
         DHA = `22:6w3`,
         DPA = `22:5w3`,
         LIN = `18:2w6`,
         SDA = `18:4w3`,
         ALA = `18:3w3`,
         ARA = `20:4w6`,
         GLA = `18:3w6`,
         sum_PUFA = sum(across(c(EPA,DHA,DPA,LIN,SDA,LIN,ARA,GLA))),
         .keep = "unused", .after = Type)

FA_groups_per <- FA|> 
  mutate(total_FA = rowSums(across(4:40)))|>
  mutate(across(4:40, ~./total_FA*100), .keep = "unused")|>
  rowwise()|>
  mutate(MUFA = sum(c_across(contains(":1"))),
         MUFA_16 = sum(c_across(contains("16:1"))),
         MUFA_18 = sum(c_across(contains("18:1"))),
         SAFA = sum(c_across(c(matches(":0"),-matches("i-")))),
         LSAFA = sum(c_across(c(`22:0`, `24:0`,`23:0`))),
         brFA = sum(c_across(contains("i-"))),
         diatom_16PUFA = sum(c_across(c(`16:2w7`,`16:3w4`))),
         green_16PUFA = sum(c_across(c(`16:3w3`,`16:4w3`,`16:2w6`))),
         w3_sum = sum(c_across(contains("w3"))),
         w6_sum = sum(c_across(contains("w6"))),
         w3_w6 = w3_sum/w6_sum,
         EPA = `20:5w3`,
         DHA = `22:6w3`,
         DPA = `22:5w3`,
         LIN = `18:2w6`,
         SDA = `18:4w3`,
         ALA = `18:3w3`,
         ARA = `20:4w6`,
         GLA = `18:3w6`,
         per_PUFA = sum(across(c(EPA,DHA,DPA,LIN,SDA,LIN,ARA,GLA))),
         .keep = "unused", .after = Type)

```

```{r FA plots}

peri_area.p <- FA_groups_areal|>
  filter(Type == "Periphyton")|>
  relocate(sum_PUFA, ARA, .before = "DPA")|>
  select(1:19,-Taxon,-Type)|>select(where(~ any(. != 0)))|>
  pivot_longer(2:17,values_to = "FA_mg.m2", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  filter(!str_detect(FA,"16PUFA")&!str_detect(FA,"_sum")&!str_detect(FA,"MUFA_")&!str_detect(FA,"brFA"))|>
  mutate(FA = fct_recode(FA,"w3:w6" = "w3_w6", "Total FA" = "Tot_FA", "PUFA"= "sum_PUFA"),
         FA = fct_reorder(FA, FA_mg.m2, .fun = median, .desc = TRUE))|>
  ggplot(aes(x = FA, y = FA_mg.m2))+
  geom_boxplot(linewidth = 0.75)+
  geom_point(size = 3, alpha = 0.5)+
  scale_y_log10(breaks = c(0.0002,0.01,0.1,1, 10, 200), labels = c("0.0002","0.01","0.1", "1", "10", "200"))+
  theme_bw()+
  ylab(bquote("Fatty Acid (mg ∙"~m^2 ~")"))+ #ALT +249 gives multiplication dot
  theme(axis.text.x = element_text(angle = 90, size = 15, face = "bold", vjust = 0.5), axis.title.x = element_blank(),
        axis.text.y = element_text(size = 15, face = "bold"), axis.title.y = element_text(size = 18, face = "bold"))
peri_area.p

peri_per.p <- FA_groups_per|>
  filter(Type == "Periphyton")|>
  relocate(ARA, per_PUFA, .before = "DPA")|>
  select(1:18,-Taxon,-Type)|>select(where(~ any(. != 0)))|>
  pivot_longer(2:16,values_to = "FA_mg.m2", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  filter(!str_detect(FA,"16PUFA")&!str_detect(FA,"_sum")&!str_detect(FA,"MUFA_")&!str_detect(FA,"brFA"))|>
  mutate(FA = fct_recode(FA,"w3:w6" = "w3_w6", "PUFA" = "per_PUFA"),
         FA = fct_reorder(FA, FA_mg.m2, .fun = median, .desc = TRUE))|>
  ggplot(aes(x = FA, y = FA_mg.m2))+
  geom_boxplot(linewidth = 0.75)+
  geom_point(size = 3, alpha = 0.5)+
  scale_y_log10(breaks = c(0.001,0.01,0.1,1,10,100), labels = c("0.001","0.01","0.1","1","10","100"))+
  theme_bw()+
  ylab(bquote("Fatty Acid (%)"))+
  theme(axis.text.x = element_text(angle = 90, size = 15, face = "bold", vjust = 0.5), axis.title.x = element_blank(),
        axis.text.y = element_text(size = 15, face = "bold"), axis.title.y = element_text(size = 18))
peri_per.p

```


```{r}
FA_groups_per|>
  filter(Type == "Periphyton")|>
  select(1:18,-Taxon,-Type)|>select(where(~ any(. != 0)))|>
  pivot_longer(2:15,values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  ggplot(aes(x = FA, y = FA_per))+geom_violin()+geom_point()+scale_y_log10()

FA_groups_per|>
  filter(Type == "Invert")|>
  select(1:18,-Taxon,-Type)|>select(where(~ any(. != 0)))|>
  pivot_longer(2:15,values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  ggplot(aes(x = FA, y = FA_per))+geom_violin()+geom_point()+scale_y_log10()
```




Hypothesis: Light and nutrients control variation in fatty acid composition.
Alt Hypothesis: Variation poorly defined by light/nutrients, instead EcoRegion or something else.
We can test this via omnibus test of a simple OLS reg model with light and nutrients predicting each FA group. We can also compare models with different predictors using AIC. 

FA groups: EPA, w3:w6, brFA, LSAFA, 

We used eta^2 as a measure of variance explained by the canopy and nutrient predictors while accounting for potentially confounding effects of EcoRegion differences that influence substrate. Although, we did not notice significant variance inflation or preliminary visualization did nto indicate correlatations among these variables. Eta^2 is the ratio of the sum of squares of a predictor divided by the total sum of squares of the model.

## Periphyton reg models
```{r areal lms}
peri_areal_lms <- FA_groups_areal|>
  filter(Type == "Periphyton")|>
  pivot_longer(c(LSAFA,ARA, Tot_FA, EPA, DHA, sum_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L, data = data)),
         model_reduced =  list(lm(FA_per ~ EcoRegion, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            Anova(model, type = 2)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            Anova(model, type = 2)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("EcoRegion"))|>summarize(pR2_ecoregion = sum(pss)),
             resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)


peri_areal_lms_outlier <- FA_groups_areal|>
  filter(Type == "Periphyton")|>
  filter(Site != "Mill")|>
  pivot_longer(c(LSAFA,ARA,Tot_FA, w3_w6,EPA, DHA, sum_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = ifelse(FA %in% c("brFA", "LSAFA"), list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L, data = data[data$Site!="Mill",])), list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L, data = data))),
         model_reduced = ifelse(FA %in% c("brFA", "LSAFA"), list(lm(FA_per ~ EcoRegion, data = data[data$Site!="Mill",])), list(lm(FA_per ~ EcoRegion, data = data))))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            Anova(model, type = 2)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)))|>
  select(-AIC,-BIC,-logLik,-deviance)

peri_areal_resids<-map(peri_areal_lms$resids, ~(ggplot(data = ., 
                                  aes(x = fitted, y = resids, shape = EcoRegion))+
                             geom_point()+
                             ggtitle(.$name)+geom_text(aes(label = Site))))

peri_areal_lms|>select(FA, r.squared, p.value)|>arrange(desc(r.squared))
peri_areal_resids

options(digits = 3)
  
areal_lm_table<-peri_areal_lms|>mutate(coef2 = map(coefs, ~select(.,term,estimate)|>pivot_wider(id_cols=everything(), 
                                                                                      values_from = "estimate", 
                                                                                      names_from = "term")))|>unnest(coef2)|>select(-df,-df.residual,-sigma,-r.squared,-nobs,-coefs,-vif,-resids)|>
  rename(`Overall R2` = "adj.r.squared",
         #`F statistic` = "statistic",
         `Overall P` = "p.value", 
          `LNH P` = "p.val_lnh",
         `LNH partial R2` = "pR2_lnh",
         Ecoregion = "EcoRegionSpringfield_Plateau",
         `Canopy Cover (%)` = "Canopy_Cover_per",
         `NO3 (ug/L)` = "NO3_ug.L")|>arrange(desc(`LNH partial R2`))

areal_lm_table_outlier<-peri_areal_lms_outlier|>mutate(coef2 = map(coefs, ~select(.,term,estimate)|>pivot_wider(id_cols=everything(), 
                                                                                      values_from = "estimate", 
                                                                                      names_from = "term")))|>unnest(coef2)|>select(-df,-df.residual,-sigma,-r.squared,-nobs,-coefs,-vif)|>
  rename(`Overall R2` = "adj.r.squared",
         #`F statistic` = "statistic",
         `Overall P` = "p.value", 
          `LNH partial P` = "p.val_lnh",
         `LNH partial R2` = "pR2_lnh",
         Ecoregion = "EcoRegionSpringfield_Plateau",
         `Canopy Cover (%)` = "Canopy_Cover_per",
         `NO3 (ug/L)` = "NO3_ug.L")|>arrange(desc(`LNH partial R2`))
                                         
                                         

areal_lm_table|>write_csv(here("analysis/paper/Table1A_PeriAreal_linmod.csv"))
#areal_lm_table_outlier|>write_csv(here("analysis/paper/TableS1A_PeriAreal_linmod-outlier.csv"))

```
LNH explained reasonable variation in the areal quantities of a few FA groups, specifically LSAFA (R^2=0.48, P= 0.011), Total FAs (R^2=0.43, P= 0.029), w3/w6 (R^2=0.34, P= 0.052), and DHA (R^2=0.30, P= 0.016). 10-20% of the variation in sumPUFAs, EPA, and ARA was explained by LNH. Examination of the residuals for LSAFA suggested one site (Mill) was an outlier and potentially driving the LNH relationship. Without this site, the LSAFA variation explained by LNH decreased to R^2 = 0.30 (P = 0.09). 


```{r percent lms}
peri_per_lms <- FA_groups_per|>
  filter(Type == "Periphyton")|>
  mutate(w3_w6 = log(w3_w6))|> #log transform because response is a ratio - Isles 202x
  #filter(Site != "Diles")|>
  pivot_longer(c(LSAFA,ARA,w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            Anova(model, type = 2)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            Anova(model, type = 2)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("EcoRegion"))|>summarize(pR2_ecoregion = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

peri_per_resids<-map(peri_per_lms$resids, ~(ggplot(data = ., 
                                  aes(x = fitted, y = resids, shape = EcoRegion))+
                             geom_point()+
                             ggtitle(.$name)+geom_text(aes(label = Site))))

peri_per_resids

per_lm_table<-peri_per_lms|>mutate(coef2 = map(coefs, ~select(.,term,estimate)|>pivot_wider(id_cols=everything(), 
                                                                                      values_from = "estimate", 
                                                                                      names_from = "term")))|>unnest(coef2)|>select(-df,-df.residual,-statistic,-sigma,-r.squared,-nobs,-coefs,-vif,-resids)|>
  rename(`Overall R2` = "adj.r.squared",
         #`F statistic` = "statistic",
         `Overall P` = "p.value", 
          `LNH P` = "p.val_lnh",
         `LNH partial R2` = "pR2_lnh",
         `Beta_Ecoregion` = "EcoRegionSpringfield_Plateau",
         `Ecoregion partial R2` = "pR2_ecoregion",
         `Beta_Canopy Cover` = "Canopy_Cover_per",
         `Beta_NO3` = "NO3_ug.L")|>arrange(desc(`LNH partial R2`))


per_lm_table|>write_csv(here("analysis/paper/Table1B_PeriPer_linmod.csv"))
```
Similar to the areal quantities, LNH explained substantial variation in %LSAFA (R^2 = 0.48, P=0.03), with partial slopes of 0.01 %LSAFA per canopy cover (percent) and 0.0015 %LSAFA per NO3 (ug/L). However, examination of residuals of the LSAFA regression suggested one site (Diles) was a strong outlier, and removal of this site strongly reduced the explanatory power of LNH for %LSAFA (R^2 = 0.20, P=0.37). LNH also explained fair variation in the w3:w6 ratio (R2 = 0.3, P = 0.05). %PUFAs were generally poorly explained by LNH. In particular, the explained variation for total %PUFA concentration and EPA were <0.03. Instead, Ecoregion explained a sizable portion of the variation in %EPA and %PUFA (R^2 = 0.5 and 0.42 respectively).






# Ordinate the periphyton FA profiles
Groups = 18 MUFA, 16_MUFA, DHA, EPA, ARA, ALA, EPA, SDA, LIN, brFA,LSAFA



RDA needs an env matrix and a response matrix
```{r peri percent RDA}
peri_group_com_per<-FA_groups_per|>
  filter(Type == "Periphyton")|>
  select(Site,LSAFA,brFA,EPA,DPA,LIN,SDA,ALA,ARA,GLA,DHA, MUFA_16, MUFA_18)

peri_env_per <- FA_groups_per|>
  filter(Type == "Periphyton")|>
  select(Canopy_Cover_per, NO3_ug.L, EcoRegion)|>
  mutate(EcoRegion = as_factor(EcoRegion))
#community matrix same as NMDS above: peri_group_com
peri_group_per <- peri_group_com_per|>select(-Site)

peri_per.rda <- rda(peri_group_per ~ EcoRegion+NO3_ug.L+Canopy_Cover_per, data = peri_env_per, scale = TRUE)

anova(peri_per.rda, by = "term",permutations = how(nperm=9999))
summary(peri_per.rda)
plot(peri_per.rda, scaling = 1,display = c("sp", "wa", "cn"))
plot(peri_per.rda, scaling = 2,display = c("sp", "wa", "cn"))

```
RDA suggested that only Ecoregion, not NO3 or Canopy Cover, strongly predicted
```{r peri areal RDA}


peri_group_com_are<-FA_groups_areal|>
  filter(Type == "Periphyton")|>
  select(Site,LSAFA,ARA,brFA,MUFA_16,MUFA_18,DHA,LIN,SDA,EPA,ALA)

peri_env_are <- FA_groups_areal|>
  filter(Type == "Periphyton")|>
  select(Canopy_Cover_per, NO3_ug.L, SRP_ug.L, EcoRegion, Hay_per)|>
  mutate(EcoRegion = as_factor(EcoRegion))
#community matrix same as NMDS above: peri_group_com
peri_group_are <- peri_group_com_are|>select(-Site)



peri_are.rda <- rda(peri_group_are ~ EcoRegion+NO3_ug.L+Canopy_Cover_per, data = peri_env_are, scale = TRUE)

anova(peri_are.rda, by = "term",permutations = how(nperm=9999))
summary(peri_are.rda)
plot(peri_are.rda, scaling = 1,display = c("sp", "wa", "cn"))
plot(peri_are.rda, scaling = 2,display = c("sp", "wa", "cn"))

```
After accounting for ecoregional differences, Canopy cover and NO3 had limited influence on periphyton fatty acid composition. This was only significant when periphyton components were expressed as percents, associating most strongly with LSAFA and ALA. When FA components were expressed as areal units, there was no overall effect.





# Inverts analysis
```{r create ratio}
FA_per_ratios<-FA_groups_per|>
    group_by(Site)|>
    mutate(across(c(LSAFA,ARA,w3_w6,EPA, DHA, per_PUFA), ~./.[Type == "Periphyton"]), .keep = "used")|>
  ungroup()

peri_FA_per <- FA_groups_per|>
  filter(Type == "Periphyton")|>
  rename_with(~str_c(.,"_peri"), .cols = c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA))|>
  select(contains(c("_peri", "Site")))

FA_per_pericols <- FA_groups_per|>
  left_join(peri_FA_per, by = "Site")

#define FFGs
ffg <- read_csv(here("analysis/data/raw_data/Ozarks_FFGs.csv"))|>
  mutate(FFG = as_factor(FFG),
         FFG = fct_collapse(FFG, 
                            `Collector-Gatherer` = c("Collect", "Gather", "Filter"),
                            Predator = "Pred",
                            Shredder = "Shred",
                            Scraper = "Scrape"))

ffg_cg <- ffg|>filter(FFG=="Collector-Gatherer")|>pull(Taxa)
ffg_shred <- ffg|>filter(FFG=="Shredder")|>pull(Taxa)
ffg_p <- ffg|>filter(FFG=="Predator")|>pull(Taxa)
ffg_sc <- ffg|>filter(FFG=="Scraper")|>pull(Taxa)

levels(ffg$FFG)

multi_ffg_taxa<-ffg|>group_by(Taxa)|>count(FFG)|>summarize(n=sum(n))|>filter(n>1)|>pull(Taxa) #only 8 observations of multi-FFG taxa

FA_groups_per|>filter(!(Taxon %in% multi_ffg_taxa))
FA_groups_per|>filter(Type == "Invert")
ffg_inverts <- FA_groups_per|>filter(!(Taxon %in% multi_ffg_taxa))|>filter(Type == "Invert")|>left_join(ffg|>rename(Taxon = "Taxa"))

```

```{r invert ffg dfs}
intersect(ffg_cg,ffg_p,ffg_shred,ffg_shred)
intersect(ffg_cg,ffg_shred)
intersect(ffg_cg,ffg_sc)
ffg_cg
ffg_p
blah<-FA_groups_per |> filter(Taxon == "Amphipod")

invert_areal_cg<-FA_groups_areal|>filter(Type == "Invert")|>filter(Taxon %in% ffg_cg)
invert_per_cg<-FA_groups_per|>filter(Type == "Invert")|>filter(Taxon %in% ffg_cg)
#invert_ratio_cg<-FA_per_ratios|>filter(Type == "Invert")|>filter(Taxon %in% ffg_cg)
invert_per_cg2<-FA_per_pericols|>filter(Type == "Invert")|>filter(Taxon %in% ffg_cg)

invert_areal_shred<-FA_groups_areal|>filter(Type == "Invert")|>filter(Taxon %in% ffg_shred)
invert_per_shred<-FA_groups_per|>filter(Type == "Invert")|>filter(Taxon %in% ffg_shred)
#invert_ratio_shred<-FA_per_ratios|>filter(Type == "Invert")|>filter(Taxon %in% ffg_shred)
invert_per_shred2<-FA_per_pericols|>filter(Type == "Invert")|>filter(Taxon %in% ffg_shred)


invert_areal_pred<-FA_groups_areal|>filter(Type == "Invert")|>filter(Taxon %in% ffg_p)
invert_per_pred<-FA_groups_per|>filter(Type == "Invert")|>filter(Taxon %in% ffg_p)
#invert_ratio_pred<-FA_per_ratios|>filter(Type == "Invert")|>filter(Taxon %in% ffg_p)
invert_per_pred2<-FA_per_pericols|>filter(Type == "Invert")|>filter(Taxon %in% ffg_p)

invert_areal_scrape<-FA_groups_areal|>filter(Type == "Invert")|>filter(Taxon %in% ffg_sc)
invert_per_scrape<-FA_groups_per|>filter(Type == "Invert")|>filter(Taxon %in% ffg_sc)
#invert_ratio_scrape<-FA_per_ratios|>filter(Type == "Invert")|>filter(Taxon %in% ffg_sc)
invert_per_scrape2<-FA_per_pericols|>filter(Type == "Invert")|>filter(Taxon %in% ffg_sc)

```

```{r correlations}
blah <- invert_per_shred |> correlation(select = c("LSAFA","ARA","ALA","LIN","brFA", "w3_w6","EPA", "DHA", "per_PUFA"), select2 = c("Canopy_Cover_per", "NO3_ug.L"))
blah
```

```{r reg model}
#mass FA/ mass dry mass
invert_mass_lms <- FA_groups_areal|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA,Tot_FA, w3_w6,EPA, DHA, sum_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA, Taxon)|>
  mutate(size = nrow(data))|>filter(size >8)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

#percent FA component of total FAs
invert_per_lms <- FA_groups_per|>
  pivot_longer(c(LSAFA,ARA,w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA, Taxon)|>
  mutate(size = nrow(data))|>filter(size >8)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)
```


```{r inverts ffgs percent lnh}
invert_scrape_lms <- invert_per_scrape|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_pred_lms <- invert_per_pred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_cg_lms <- invert_per_cg|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_shred_lms <- invert_per_shred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+Canopy_Cover_per+NO3_ug.L+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

```
```{r inverts ffgs percent afdm and chla}
invert_scrape_afdmlms <- invert_per_scrape|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+afdm_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_pred_afdmlms <- invert_per_pred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+afdm_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_cg_afdmlms <- invert_per_cg|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+afdm_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_shred_afdmlms <- invert_per_shred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+afdm_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_scrape_chllms <- invert_per_scrape|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_pred_chllms <- invert_per_pred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_cg_chllms <- invert_per_cg|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_shred_chllms <- invert_per_shred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

```

```{r invert ffg}

invert_scrape_chllms <- invert_per_scrape|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA,w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_pred_chllms <- invert_per_pred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_cg_chllms <- invert_per_cg|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)

invert_shred_chllms <- invert_per_shred|>
  filter(Type == "Invert")|>
  pivot_longer(c(LSAFA,ARA, w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lm(FA_per ~ EcoRegion+chla_mg.m2+Taxon, data = data)),
         model_reduced = list(lm(FA_per ~ EcoRegion+Taxon, data = data)))|>
  summarize(glance(model), 
            coefs = list(tidy(model)|>filter(term != "(Intercept)")), 
            vif = list(tibble(vif(model))),
            p.val_lnh= anova(model,model_reduced)[[6]][[2]],
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )))|>
  select(-AIC,-BIC,-logLik,-deviance)
```


```{r invert lmm attempt}

lmm1 <- ffg_inverts|>
  pivot_longer(c(LSAFA,ARA,w3_w6,EPA, DHA, per_PUFA),values_to = "FA_per", names_to = "FA")|>mutate(FA = as_factor(FA))|>
  nest_by(FA)|>
  mutate(model = list(lmer(FA_per ~ Canopy_Cover_per*FFG+NO3_ug.L*FFG+(1|FFG/Taxon), data = data)),
         model_lm = list(lm(FA_per ~ Canopy_Cover_per*FFG+NO3_ug.L*FFG, data = data)),
         model_reduced = list(lmer(FA_per ~ 1 + FFG + (1|Taxon), data = data)))|>
  summarize(glance(model), 
            coefs_lmm = list(tidy(model)|>filter(term != "(Intercept)")), 
            coefs_lm = list(tidy(model_lm)|>filter(term != "(Intercept)")),
            #vif = list(tibble(vif(model))),
           comparison = list(anova(model,model_reduced)),
            anova(model)|>as_tibble(rownames = "term")|>rename(ss= `Sum Sq`)|>mutate(sum_ss = sum(ss), pss = ss/sum_ss)|>filter(term %in% c("Canopy_Cover_per","NO3_ug.L"))|>summarize(pR2_lnh = sum(pss)),
            resids = list(bind_cols(data, 
                                    resids = resid(model),
                                    fitted = fitted(model), name = FA, 
                                    )),
           predicted_values_lmm = list(modelr::data_grid(data, Canopy_Cover_per, NO3_ug.L,nesting(Taxon,FFG))|>modelr::add_predictions(model)|>rename("{FA}" := "pred")),
           predicted_values_lm = list(modelr::data_grid(data, Canopy_Cover_per, NO3_ug.L,FFG)|>modelr::add_predictions(model_lm)|>rename("{FA}" := "pred")))|>
  select(-AIC,-BIC,-logLik)

ggplot()+
  geom_smooth(data = lmm1[[11]][[4]], aes(y = EPA, x = Canopy_Cover_per, group = FFG), color = "black",se = FALSE)+
  geom_smooth(data = lmm1[[11]][[4]], aes(y = EPA, x = Canopy_Cover_per, color = Taxon), se = FALSE)+
  geom_point(data = ffg_inverts,aes(y = EPA,x = Canopy_Cover_per, color = Taxon, shape = FFG))+
  facet_wrap(.~FFG, scales = "free")

ggplot()+
  geom_smooth(data = lmm1[[11]][[4]], aes(y = EPA, x = NO3_ug.L, group = FFG),color = "black", se = FALSE)+
  geom_smooth(data = lmm1[[11]][[4]], aes(y = EPA, x = NO3_ug.L, color = Taxon), se = FALSE)+
  geom_point(data = ffg_inverts,aes(y = EPA,x = NO3_ug.L, color = Taxon, shape = FFG))+
  facet_wrap(.~FFG, scales = "free")

ggplot()+
  geom_smooth(data = lmm1[[12]][[4]], aes(y = EPA, x = Canopy_Cover_per, group = FFG), se = FALSE)+
  geom_point(data = ffg_inverts,aes(y = EPA,x = Canopy_Cover_per, color = FFG))+
  facet_wrap(.~FFG, scales = "free")
```

Remember it's worth just modeling the intercepts!!! That variation among groups and taxa is very interesting
```{r pull emmeans lmm}
#library(emmeans)

m_test <- lmer(EPA ~ Canopy_Cover_per*FFG+(1|FFG/Taxon), data = ffg_inverts)
summary(m_test)
m.trend<-emtrends(m_test, specs = "FFG", var = "Canopy_Cover_per", addl.vars = "Taxon")
contrast(m.trend)

contrast(m.grid)|>tidy()
plot(m.grid)

```

# Inverts and Peri - try Procrustes a la Guo "dark side" paper







# Old







```{r}
PUFA_invert <- FA  |> group_by(Taxon)%>%
         filter(n()>2&Type == "Invert")|>mutate(Taxon = as_factor(Taxon), Site = as_factor(Site))|>
  select(where(~ any(. != 0)))|>ungroup()

PUFA_peri <- FA |>
  filter(Type == "Periphyton")|>
  mutate(across(all_of(4:60), ~.*Slurry_Volume_mL/1000/Cobble_SA_cm2))

PUFA_peri_afdm <- FA |>
  filter(Type == "Periphyton")|>
  mutate(across(all_of(4:60), ~.*Slurry_Volume_mL/Slurry_AFDM_g)) #will be ng FA/per mg AFDM, in accordance with invert values

PUFA_peri_per <- FA |>
  filter(Type == "Periphyton")|>
  mutate(total_FA = rowSums(across(4:60)))|>
  mutate(across(all_of(4:60), ~./total_FA*100), .keep = "unused")
  


invert_pufa_mat <- PUFA_invert |> select(`18:3w3`,`18:4w3`,`20:5w3`, `18:2w6`, `20:4w6`,`22:5w3`,`22:6w3`)
invert_exp_mat <- PUFA_invert |> select(-(4:39))

invert_pufa_mat_phys <- PUFA_invert|>filter(Taxon != "Physidae") |> select(`18:3w3`,`18:4w3`,`20:5w3`, `18:2w6`, `20:4w6`,`22:5w3`,`22:6w3`)
invert_exp_mat_phys <- PUFA_invert |>filter(Taxon != "Physidae") |> select(-(4:39))

peri_pufa_mat <- PUFA_peri|> select(`18:3w3`,`18:4w3`,`20:5w3`,`22:5w3`,`22:6w3`, `18:2w6`, `20:2w6`, `20:4w6`)
peri_pufa_per_mat <- PUFA_peri_per|> select(4:60)|>select(where(~any(. != 0)))|>
  select(`18:3w3`,`18:4w3`,`20:5w3`,`22:5w3`,`22:6w3`, `18:2w6`, `20:2w6`, `20:4w6`)
peri_exp_mat <- PUFA_peri|> select(-(4:60))

ggplot(FA, aes(x = EcoRegion, y = NO3_ug.L))+geom_boxplot()+geom_point()
ggplot(FA, aes(x = EcoRegion, y = Canopy_Cover_per))+geom_boxplot()+geom_point()
ggplot(FA, aes(NO3_ug.L, Canopy_Cover_per))+geom_label(aes(label = Site))

#Mill has very high pasture cover
```


# Periphyton analysis
For periphyton, we want to analyze whether PUFA composition is driven mostly by 

## Hypothesis testing
Want to test hypothesis that light and/or nutrients drive PUFA at % or per unit stream bottom

-PUFA percent models

-PUFA areal mass models



## Exploratory analysis
PC1 explains like 80% of the variation in peri PUFAs - could regress site scores on PC1 against different environmental variables 


```{r periphyton pca}
pc_pufa <- prcomp(peri_pufa_per_mat,  scale = TRUE)
autoplot(pc_pufa, label = TRUE, loadings = TRUE, data = PUFA_peri, loadings.label = TRUE, 
         label.label = 'Site', label.size = 2.5, label.repel = TRUE, xlim = c(-0.9,0.9), shape = "EcoRegion", size = 5, alpha = 0.5, main = "Periphyton PUFA PCA")+theme_bw()
pc.s <- scores(pc_pufa, 1:2, display = "species")|>as_tibble()|>mutate(names = colnames(peri_pufa_per_mat))
pc.loadings <- pc_pufa$rotation

blah <- PUFA_peri|>bind_cols(pc.s)

ggplot(blah, aes(y = PC2, x = ForTot, size = Canopy_Cover_per, shape = EcoRegion))+geom_point()
```

## RDA Models
### Local predictors
```{r periphyton  per}
peri_exp_mat_preds <- peri_exp_mat|>select(Canopy_Cover_per,SRP_ug.L, NO3_ug.L,EcoRegion, UrbTot_per, Hay_per)

upr <- rda(peri_pufa_per_mat ~ ., data = peri_exp_mat_preds, scale= FALSE)
lwr <- rda(peri_pufa_per_mat ~ 1, data = peri_exp_mat_preds, scale= FALSE)
peri_local_mods <- ordistep(lwr, direction = "both", scope = formula(upr))

#peri_rda_comp <- rda(peri_pufa_mat ~ SRP_ug.L, data = peri_exp_mat, scale = TRUE)


anova(peri_local_mods, by = "term",permutations = how(nperm=9999))
summary(peri_local_mods)
plot(peri_local_mods, scaling = 2)
```

```{r periphyton  per SA}
peri_exp_mat_preds <- peri_exp_mat|>select(UrbTot_per,Canopy_Cover_per,SRP_ug.L, NO3_ug.L,EcoRegion, Hay_per)

upr <- rda(peri_pufa_mat ~ ., data = peri_exp_mat_preds, scale= TRUE)
lwr <- rda(peri_pufa_mat ~ 1, data = peri_exp_mat_preds, scale= TRUE)
peri_local_mods <- ordistep(lwr, direction = "both", scope = formula(upr))

anova(peri_local_mods, by = "axis",permutations = how(nperm=9999))
summary(peri_local_mods)
plot(peri_local_mods, scaling = 1)
```
### Watershed predictors
```{r periphyton watershed}
peri_exp_mat_ws <- peri_exp_mat|>select(EcoRegion,ForTot, UrbTot_per, Hay_per, NO3_ug.L,SRP_ug.L,Canopy_Cover_per)

upr <- rda(peri_pufa_mat ~ EcoRegion*Canopy_Cover_per, data = peri_exp_mat_ws, scale = TRUE,permutations = how(nperm=9999))
lwr <- rda(peri_pufa_mat ~ 1, data = peri_exp_mat_ws, scale = TRUE,permutations = how(nperm=9999))
peri_ws_mods <- ordistep(lwr, scope = formula(upr))
anova(peri_ws_mods, by = "margin", permutations = how(nperm=9999))
plot(peri_ws_mods, scaling = "species", display = c("sp", "bp"))

peri_rda_urb<-rda(peri_pufa_mat ~ ForTot, data = peri_exp_mat, scale = TRUE)
peri_rda_urb
peri_rda_for
peri_rda_for
peri_rda_pas
peri_rda_pas

anova(peri_ws_mods, by = "margin",permutations = how(nperm=9999))

ggplot(data = peri_exp_mat, aes(x=UrbTot_per, y = SRP_ug.L))+geom_point()


```


### Cross-scale predictors


# Invert models
## PCA visualization
```{r invert pca}
ggplot(data = PUFA_invert|>filter(Taxon == "Psephenidae"), 
       aes(x = Canopy_Cover_per, y = `20:5w3`))+geom_point()

levels(PUFA_invert$Taxon)
pc_pufa <- prcomp(invert_pufa_mat,  scale = TRUE)
autoplot(pc_pufa, loadings = TRUE, data = PUFA_invert, loadings.label = TRUE, 
          label.size = 3, size=5, alpha = 0.5, shape = 'EcoRegion', main = "Invert PUFA PCA")+theme_bw()

pc_pufa <- prcomp(invert_pufa_mat_phys,  scale = TRUE)
autoplot(pc_pufa, loadings = TRUE, data = PUFA_invert|>filter(Taxon != "Physidae"), loadings.label = TRUE, 
          label.size = 3, size=5, alpha = 0.5, shape = 'EcoRegion', main = "Invert PUFA PCA")+theme_bw()

autoplot(pc_pufa, label = TRUE, loadings = TRUE, data = FA_invert, loadings.label = TRUE, 
         label.label = 'Site', label.size = 3,  label.repel = TRUE, ylim = c(-0.1,0.1), main = "Invert PUFA PCA")+theme_bw()
```

## RDA Models

### Local predictors
```{r invert local}
invert_rda_can
invert_rda_can
invert_rda_chla
invert_rda_chla
invert_rda_srp
invert_rda_srp
invert_rda_no3
invert_rda_no3


```

### Watershed predictors
```{r invert watershed}
invert_exp_mat_ws <- invert_exp_mat_phys|>select(EcoRegion,Taxon,ForTot, UrbTot_per, Hay_per, NO3_ug.L,SRP_ug.L,Canopy_Cover_per, chla_mg.m2)

upr <- rda(invert_pufa_mat_phys ~ Taxon*EcoRegion, data = invert_exp_mat_ws, scale = TRUE,permutations = how(nperm=9999))
lwr <- rda(invert_pufa_mat_phys ~ 1, data = invert_exp_mat_ws, scale = TRUE,permutations = how(nperm=9999))
invert_ws_mods <- ordistep(lwr, scope = formula(upr))
anova(invert_ws_mods, by = "margin", permutations = how(nperm=9999))
plot(invert_ws_mods, scaling = "species", display = c("sp", "bp"))

peri_rda_urb
peri_rda_urb
peri_rda_for
peri_rda_for
peri_rda_pas
peri_rda_pas

```
```{r}
upr <- rda(invert_pufa_mat ~ Hay_per*Canopy_Cover_per+Condition(Taxon), data = invert_exp_mat_ws, scale = TRUE,permutations = how(nperm=9999))
lwr <- rda(invert_pufa_mat ~ 1, data = invert_exp_mat_ws, scale = TRUE,permutations = how(nperm=9999))
invert_ws_mods <- ordistep(lwr, scope = formula(upr))

```

### Join

```{r}
#need to make sure these are apples to oranges comparisons of PUFA dissimilarities

peri.1 <- PUFA_peri_afdm |>select(Canopy_Cover_per,EcoRegion,Site, `18:3w3`,`18:4w3`,`20:5w3`, `18:2w6`, `18:2w6`, `20:4w6`,`22:6w3`)|>
  rename(peri_ara = "20:4w6", 
         peri_lin = "18:2w6",
         peri_gla = "18:2w6",
         peri_ala = "18:3w3", 
         peri_sda = "18:4w3", 
         peri_epa = "20:5w3", 
         peri_dha = "22:6w3")
  
invert.1<-PUFA_invert|>select(Canopy_Cover_per,EcoRegion,Site, Taxon, `18:3w3`,`18:4w3`,`20:5w3`, `18:2w6`,`18:2w6`, `20:4w6`,`22:6w3`)|> 
  rename(ara = "20:4w6", 
         lin = "18:2w6",
         gla = "18:2w6",
         ala = "18:3w3", 
         sda = "18:4w3", 
         epa = "20:5w3", 
         dha = "22:6w3")

PUFA_join <- left_join(invert.1, peri.1, by = c("Canopy_Cover_per","EcoRegion","Site"))|>
  mutate(
ara_r = ara/peri_ara,
lin_r = lin/peri_lin,
gla_r = gla/peri_gla,
ala_r = ala/peri_ala,
sda_r = sda/peri_sda,
epa_r = epa/peri_epa,
dha_r = dha/peri_dha)
#by = c("18:3w3","18:4w3","20:5w3", "18:2w6", "20:4w6","22:5w3","22:6w3")


ggplot(data = PUFA_join, aes(x = peri_ara, y = ara, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_lin, y = lin, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_gla, y = gla, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_ala, y = ala, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_sda, y = sda, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_epa, y = epa, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_dha, y = dha, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")

ggplot(data = PUFA_join, aes(x = EcoRegion, y = ara, color = Taxon))+geom_boxplot()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_lin, y = lin, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_gla, y = gla, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_ala, y = ala, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_sda, y = sda, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = Canopy_Cover_per, y = epa_r, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")
ggplot(data = PUFA_join, aes(x = peri_dha, y = dha, color = Taxon))+geom_point()+facet_wrap(.~Taxon, scales = "free")

```
