---
title: "Ozarks_SIA_Analysis"
author: "CJR"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(readxl)
library(datawizard)

```


```{r}

sia_master<-read_xlsx(here("analysis/data/raw_data/Ozarks_SIA_master.xlsx"))%>%
  select(Site, Type, Taxon, Sample, d15N,d13C, C_per,N_per,CN_mass_ratio)%>%
  filter(Site != "BLANK")%>%
  mutate(Acidified = as_factor(ifelse(str_detect(Type, "acid"), "Yes", "No")))%>%
  mutate(Type = as_factor(str_remove(Type,"-acid")))%>%
  mutate(Site = fct_collapse(Site,
                             `Cap Fork-North Sylamore` = c("Cap Fork", "Capfork"),
                             `Cove` = c("Cove", "Cove Creek"),
                             `East Livingstone` = c("E. Livingston", "E. Livingstone"),
                             `Lewis Prong-Panther` = c("Lewis Prong", "Lewis Prong Panther"),
                             `North Sylamore` = c("Sylvamore", "Sylamore"),
                             `Lick Branch-Little Mulberry` = "Lick Branch",
                             Spring = "Spring Creek"))%>%
  mutate(Taxon = fct_collapse(Taxon, 
                              Chloroperlidae = "ChloroPerlidae",
                              Heptageniidae = "Heptageniid",
                              Hydropsychidae = c("Hydropsychid","Hydropsyhid", "Hydrospychid", "Hydrospychidae"),
                              Corydalidae_Lg = c("Large Megaloptera", "Large_Megaloptera"),
                              Corydalidae_sm = c("Small Megaloptera", "Small_Megaloptera"),
                              Corydalidae_Lg = "Hellgrammite",
                              Heptageniidae = "Heptgeniid",
                              Oligochaete = "Oligachaete",
                              Physidae = "Physidaes",
                              Psephenidae = "Psephenis",
                              Tabanidae = c("Tabinid", "Tabanid"),
                              Tipulidae = "Tipulid",
                              `NA` = c("N/A", "NA")
                              ),
         Taxon = na_if(Taxon, "NA"),
         Taxon = fct_drop(Taxon,"NA"))




```


```{r}

sia2<- sia_master%>%
  select(-Sample)%>%
  group_by(Site, Type,Taxon,Acidified)%>%
  summarize(d13C_uncorr = mean(d13C),
            d15N = mean(d15N),
            C_per_uncorr = mean(C_per),
            N_per = mean(N_per))

sia_acids <- sia2%>%
  filter(Acidified == "Yes")%>%
  rename(d13C_acid = d13C_uncorr,
         C_per_acid = C_per_uncorr)%>%
  select(Site, Type, Taxon, d13C_acid, C_per_acid)

sia <- sia2%>%
  filter(Acidified == "No")%>%
  select(-Acidified)%>%
  full_join(sia_acids, by = c("Site", "Type", "Taxon"))%>%
  ungroup()%>%
  mutate(d13C = ifelse(is.na(d13C_acid), d13C_uncorr, d13C_acid),
         C_per = ifelse(is.na(C_per_acid), C_per_uncorr, C_per_acid))%>%
  mutate(CN_molar = (C_per/N_per)*(14/12),
         con_res = fct_collapse(Type, 
                                Resource = c("Biofilm", "Algae", "CPOM", "FPOM"),
                                other_level = "Consumer"))



```

```{r sia view}
#distributions of resource values
sia_distros <- sia%>%
  select(d13C,d15N,N_per,C_per,CN_molar)%>%
  describe_distribution()

sia_distros

#simple mixing models

```

```{r}
sia%>%
  filter(Type != "FPOM")%>%
  ggplot(aes(x = d13C, y = d15N, color = con_res, shape = Type))+
  geom_point()+
  facet_wrap(.~Site)

```
Note: Little Piney periphyton was lost by the time it was needed to analyze for SIA, so algae are not included here
-Mill, Lewis prong, Mulberry missing d15N values for separated algae



simmr

Source data need to have sds but target data need to be individuals. May need to consider checking some physid data that have low d13C and low %C - possible carbonate contamination?


```{r create source data}

sia2<- sia_master%>%
  select(-Sample)%>%
  filter(Type != "FPOM")%>%
  group_by(Site, Type,Taxon,Acidified)%>%
  summarize(d13C_uncorr_mean = mean(d13C, na.rm = TRUE),
            C_per_uncorr = mean(C_per, na.rm = TRUE),
            N_per = mean(N_per, na.rm = TRUE),
            d15N_mean = mean(d15N, na.rm = TRUE),
            d13C_uncorr_sd = sd(d13C, na.rm = TRUE),
            d15N_sd = sd(d15N, na.rm = TRUE))

sia_acids <- sia2%>%
  filter(Acidified == "Yes")%>%
  rename(d13C_acid = d13C_uncorr_mean,
         C_per_acid = C_per_uncorr,
         d13C_acid_sd = d13C_uncorr_sd)%>%
  select(Site, Type, Taxon, d13C_acid, C_per_acid,d13C_acid_sd)

sia_sources <- sia2%>%
  filter(Acidified == "No")%>%
  select(-Acidified)%>%
  full_join(sia_acids, by = c("Site", "Type", "Taxon"))%>%
  ungroup()%>%
  mutate(d13C_mean = ifelse(is.na(d13C_acid), d13C_uncorr_mean, d13C_acid),
         d13C_sd = ifelse(is.na(d13C_acid), d13C_uncorr_sd, d13C_acid_sd),
         C_per = ifelse(is.na(C_per_acid), C_per_uncorr, C_per_acid))%>%
  mutate(CN_molar = (C_per/N_per)*(14/12),
         con_res = fct_collapse(Type, 
                                Resource = c("Biofilm", "Algae", "CPOM", "FPOM"),
                                other_level = "Consumer"))%>%
  filter(Type != "INVERT", Type != "Biofilm")%>%
  #filter(Type != "INVERT")%>%
  select(Site, Type, d13C_mean, d15N_mean, d13C_sd, d15N_sd)





```



```{r}

#none of the inverts needed acidification
sia_targets <- sia_master%>%
  filter(Type == "INVERT")%>%
  select(Site, Taxon, d13C, d15N, C_per, N_per)%>%
  filter(!is.na(d13C))%>%
  mutate(Taxon = as_factor(Taxon))

```

```{r filter target and source to one site}
sia_targets_CF <- sia_targets%>%
  filter(Site == "Cap Fork-North Sylamore")|>
  mutate(Taxon = as.character(Taxon))
sia_targets_CF$Taxon

sia_sources_CF <- sia_sources%>%
  filter(Site == "Cap Fork-North Sylamore")

sia_targets_CF <- sia_targets%>%
  filter(Site == "Roasting Ear")|>
  mutate(Taxon = as.character(Taxon))
sia_targets_CF$Taxon

sia_sources_CF <- sia_sources%>%
  filter(Site == "Roasting Ear")

```






```{r simmr test load}
CF_simmr = simmr_load(mixtures = sia_targets_CF[, 3:4],
                         source_names = as.character(sia_sources_CF$Type),
                         source_means = as.matrix(sia_sources_CF[,3:4]),
                         source_sds = as.matrix(sia_sources_CF[,5:6]),
                        # correction_means = as.matrix(TEFs[,2:3]),
                        # correction_sds = as.matrix(TEFs[,4:5]),
                        # concentration_means = as.matrix(concdep[,2:3]),
                         group = sia_targets_CF$Taxon
                      )

```


```{r simmr test viz}
plot(CF_simmr)

```
```{r}
simmr_out_cf <- simmr_mcmc(CF_simmr)

summary(simmr_out_cf, type = "diagnostics")
post_pred_cf <- posterior_predictive(simmr_out_cf)
post_pred_cf
```
```{r}
summary(simmr_out_cf, type = "quantiles", group = 2)
plot(simmr_out_cf,type = "matrix")
compare_groups(simmr_out_cf, source = "Algae", groups = seq(1:length(unique(sia_targets_CF$Taxon))))
```

For sites with no d15N data for algae - calculate %algal contribution using d13C only, then add to results dataframe
